---
title: "Project X"
---

Welcome to my/our project page.

Keep an eye on this space to stay updated with my project activities.

(The titles below are provided as examples; please feel free to adjust them as necessary.)

# 1. Project Overview and Scope

xxxxxx

# 2. Data

Global Fuel Prices & Economic Indicators (2020-2026)

The global energy landscape has faced significant volatility between 2020 and 2026. This dataset provides a standardized, weekly record of retail fuel prices across 84 countries, aligned with global crude oil benchmarks and national fiscal indicators. It serves as a unified resource for analyzing the relationship between international commodity markets and local retail costs.

We imported our dataset directly from Kaggle.com into RStudio and began the analysis. Before using the data, we inspected the columns to ensure that there were no missing values (NAs) or data quality issues. The dataset was already in a ready-to-use format, so no additional data cleaning or preprocessing was required.

## 2.1 Data Source

[Click for Data](https://www.kaggle.com/datasets/belbino/global-fuel-prices-20202026/data){.btn .btn-dark}

## 2.2 General Information About Data

This dataset contains 27468 observations tracking the prices of Petrol, Diesel, and LPG. Each entry is enriched with regional, economic, and policy-level metadata to provide a holistic view of the global fuel economy.

Temporal Coverage: Weekly data from January 2020 to April 2026.

Global Scale: 84 countries across 7 geographic regions.

Economic Context: Integrated Brent Crude benchmarks, tax percentages, and subsidy levels.

Normalization: All pricing data is standardized to USD per Liter.


## 2.3 Reason of Choice

This dataset was chosen because fuel prices are highly sensitive to global geopolitical events, particularly wars and conflicts. Recent events such as the ABD–Iran war have caused significant volatility in oil supply and Brent crude prices, leading to fluctuations in petrol, diesel, and LPG prices worldwide. Therefore, the dataset is suitable for analyzing how geopolitical instability affects fuel prices across countries, including Turkey.

This study also compares the impact of major geopolitical events, specifically the Russia–Ukraine conflict in February 2022 and the United States–Iran conflict in March 2026, on fuel prices across different countries. The analysis focuses on differences in price sensitivity between countries, with a particular emphasis on Turkey. The aim is to assess how strongly fuel prices in Turkey are affected compared to other countries such as the United States during these periods.

Additionally, we aim to analyze the trend of fuel prices in Turkey between 2020 and 2026, examining how prices responded to the specified geopolitical events and assessing the nature of these responses over time.

## 2.4 Preprocessing

xxxxxx

# 3. Analysis

xxxxxx

## 3.1 Exploratory Data Analysis

xxxxxx

## 3.2 Trend Analysis

xxxxxx

## 3.3 Model Fitting

xxxxxx

## 3.4 Results

xxxxxx

# 4. Results and Key Takeaways

```{r}
library(tidyverse)
```
global_fuel_prices<-read.csv("/Users/seymasezer/Desktop/emu660-spring2026-seymasezery/global_fuel_prices_2020_2026.csv")
```{r}
global_fuel_prices<-read.csv("/Users/buseerdogan/Desktop/EMU660/emu660-spring2026-buseyidogan26/global_fuel_prices_2020_2026.csv")
```

```{r}
str(global_fuel_prices)
```

```{r}
ncol(global_fuel_prices)
```

```{r}
nrow(global_fuel_prices)
```

```{r}
colSums(is.na(global_fuel_prices))
```
```{r}
#Petrol zaman trendi
global_fuel_prices|>ggplot(aes(x=date, y=petrol_usd_liter))+geom_line(color="blue")+xlab("Date")+ylab("USD per Liter")+ggtitle("Petrol Price Over Time")
```

```{r}
#Diesel zaman trendi
global_fuel_prices|>ggplot(aes(x=date, y=diesel_usd_liter))+geom_line(color="red")+xlab("Date")+ylab("USD per Liter")+ggtitle("Diesel Price Over Time")
```

```{r}
#LPG zaman trendi
global_fuel_prices|>ggplot(aes(x=date, y=lpg_usd_liter))+geom_line(color="green")+xlab("Date")+ylab("USD per Liter")+ggtitle("LPG Price Over Time")
```

```{r}
global_fuel_prices |>
  ggplot(aes(x = country, y = petrol_usd_liter)) +
  geom_boxplot(fill = "lightblue") +
  geom_point(position = position_jitter(width = 0.2), alpha = 0.4) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("country")+ylab("USD per Liter")+ggtitle("Petrol Prices by Country")
```
```{r}
top_countries <- global_fuel_prices |>
  group_by(country) |>
  summarise(avg = mean(petrol_usd_liter, na.rm = TRUE)) |>
  arrange(desc(avg)) |>
  slice_head(n = 10) |>
  pull(country)
```

```{r}
global_fuel_prices|>filter(country %in% top_countries)|>
  ggplot(aes(x = country, y = petrol_usd_liter)) +
  geom_boxplot(fill = "lightblue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("country")+ylab("USD per Liter")+ggtitle("Petrol Prices by Country")
```


