---
title: "Assignment 1"
---

## (a)

## (b)

## (c)

### Data Check

```{r}
data("mtcars")
head(mtcars)
```

### Custom Summary Function

The function `compute_stats` calculates several descriptive statistics **for a numeric vector.**

It returns the mean, median, variance, interquartile range (IQR), minimum, and maximum values as a **named list.**

```{r}
compute_stats <- function(x) {
  list(
    mean = mean(x, na.rm = TRUE),
    median = median(x, na.rm = TRUE),
    variance = var(x, na.rm = TRUE),
    IQR = IQR(x, na.rm = TRUE),
    min = min(x, na.rm = TRUE),
    max = max(x, na.rm = TRUE)
  )
}
```

### Applying the Function Using a Loop:

In this section, the `compute_stats` function is applied to each numeric column of the `mtcars` dataset using a for loop. The computed statistics for each column are printed along with the column name.

```{r}
names(mtcars) #to see all column names
```

In this section, a `for` loop is used to iterate over all numerical columns of the `mtcars` dataset. For each column, the `compute_stats` function is applied to calculate previous summary statistics. The computed statistics are printed for each column along with the corresponding column name.

```{r}
for (column_name in names(mtcars)) {

  column_data <- mtcars[[column_name]]

  if (is.numeric(column_data)) {

    stats <- compute_stats(column_data)

    cat("Statistics for column:", column_name, "\n")
    print(stats)
  }
}
```

Since the output is a little bit long, 'data.frame' is tried below.

```{r}
results <- list()

for (column_name in names(mtcars)) {

  column_data <- mtcars[[column_name]]
  
  stats <- compute_stats(column_data)
  
  results[[column_name]] <- unlist(stats)
}

stats_table <- as.data.frame(do.call(rbind, results))

stats_table
```

### Applying Sapply Function

In this section, the `sapply` function is used as an alternative to a `for` loop to apply the `compute_stats` function across all numerical columns of the `mtcars` dataset. The function is automatically applied to each column, and the resulting statistics are collected in a structured output.

With this function, the output appears similar to a data frame structure.

```{r}
sapply(mtcars, compute_stats)
```

### Applying Apply Function

In this section, the `apply` function is used to apply `compute_stats` to all numeric columns of the `mtcars` dataset.

Since `apply` works on matrices, the numeric subset of the dataset is converted to a matrix.

The function is applied column-wise (margin = 2), producing summary statistics for each column.

```{r}
#Choosing only numeric columns
numeric_cols <- mtcars[, sapply(mtcars, is.numeric)]

stats_apply <- apply(numeric_cols, 2, compute_stats)

stats_apply

```

## (d)

### Polls US Election 2016

```{r}
library(dslabs)
data("polls_us_election_2016")
head(polls_us_election_2016)
```

```{r}
my_first <- "Buse"
my_birth_year <- 2000
k <- (nchar(my_first) + my_birth_year) %% 15 + 8
print(k)
```

```{r}
my_first <- "Buse"
my_birth_year <- 2000
k <- (nchar(my_first) + my_birth_year) %% 15 + 8

if (k %% 2 == 0) {
  head(polls_us_election_2016, k)
} else {
  tail(polls_us_election_2016, k)
} #since my k value is odd number, else is returned.
```

First, the total number of missing values (NA) in the entire dataset is calculated.

Next, the number of missing values in each column is calculated. The columns are sorted from the highest number of missing values to the lowest, and the top 8 columns are displayed.

```{r}
# Total NA
sum(is.na(polls_us_election_2016))

sort(colSums(is.na(polls_us_election_2016)), decreasing = TRUE)[1:8]
```

In this section, all missing values in the dataset are replaced according to the specified rules.

For numeric columns, missing values are replaced with `my_birth_year + k`. For character or factor columns, missing values are replaced with a string created using `paste0(my_first, "_", k)`.

The modified dataset is stored as a new object called `new_data`, leaving the original dataset unchanged.

```{r}
new_data <- polls_us_election_2016

for(col in names(new_data)) {
  
  if(is.numeric(new_data[[col]])) {
    
    new_data[[col]][is.na(new_data[[col]])] <- my_birth_year + k
    
  } else if(is.factor(new_data[[col]])) {
    
    new_data[[col]] <- as.character(new_data[[col]])
    new_data[[col]][is.na(new_data[[col]])] <- paste0(my_first, "_", k)
    
  } else if(is.character(new_data[[col]])) {
    
    new_data[[col]][is.na(new_data[[col]])] <- paste0(my_first, "_", k)
    
  }
}
```

```{r}
head(new_data)
```

Displaying 17 rows of the new data frame:

```{r}
my_first <- "Buse"
my_birth_year <- 2000
k <- (nchar(my_first) + my_birth_year) %% 15 + 8

if (k %% 2 == 0) {
  head(new_data, k)
} else {
  tail(new_data, k)
}
```

```{r}
sum(is.na(new_data))
```

```{r}
anyNA(new_data)
```

### AI Usage Statement

ChatGPT was consulted for guidance on implementing loops, using functions such as \`sapply\` and \`apply\`, and handling missing values in data frames.\
\
All outputs produced by the AI tool were carefully reviewed, tested, and modified when necessary using my own understanding to ensure correctness.\
\
Example prompts used include:\
\
- "How can I apply a custom function to all columns of the mtcars dataset using a for loop in R?"\
- "How can I replace NA values differently for numeric and character columns in a data frame?"\
- "How can I count the number of NA values per column in an R dataset?"
